---
title: "7313 Take-Home Exam"
output: pdf_document
## no author, anonymous grading
---

```{r, include = FALSE, message = FALSE}
# Load libraries
library(RMySQL) 
library(caret)
library(randomForest)
library(MLmetrics)
library(ranger)
library(tidyverse)
```

```{r sql_con, include = FALSE}
#Notice the naming of the code chunks, these are used for debugging
con = dbConnect(MySQL(), dbname = "BnS",
                host = "mysql-1.cda.hhs.se", port = 3306,
                user = "bns", password = "bns@sse")
```

# Importing from SQL
```{r import, message = FALSE, warning = FALSE}
df = dbGetQuery(con, 
  "SELECT 
    receipt_id,
    amount,
    quantity,
    MONTH(purchase_date) AS month,
    age,
    m_status,
    gender,
    enrollment,
    MAX(is_online) AS is_online
  FROM Transactions
    LEFT JOIN Customers
    USING (customer_id)
    LEFT JOIN Unseen 
    USING (receipt_id)
  WHERE id42205 = 1 ##change to your studentid
    OR is_online IS NOT NULL
  GROUP BY receipt_id")
```

```{r, include=F}
lapply( dbListConnections( dbDriver( drv = "MySQL")), dbDisconnect)
```

# Data preparation
## I decided to include the variables you can see in the MySQL import mostly for plausability reasons. The customer variable 'created' has no straight forward relationship with 'is_online'. The subscription variables ('no_contact', 'no_email', etc.) also can't plausibly predict whether a transaction is online; there is no causal rationale. The model might fit well to the values of these variables, but I would question external validity questionable then. Originally, I included the variable 'city'. There are 365,126 missing value for 'city'. This is 85.30% of all values. I ignore this variable, since no appropriate imputation can be determined (the mode is unlikely to deliver a representative result).

```{r NAs}
for(i in 1:ncol(df)) {print(paste(colnames(df)[i], "has", toString(sum(is.na(df[, i]))), "missing values."))}
```

## 1 missing value for age. 1 missing value for m_status. 1 missing value for gender. Looking at the data, it becomes apparent that 'customer_id' in Transactions has been miscoded in one instance. It is easiest to remove the respective observation. Now, the categorical variables have to be turned into factors.
```{r factors}
df = df[!is.na(df$age),]

df$m_status = as.factor(df$m_status)
df$gender = as.factor(df$gender)
df$enrollment = as.factor(df$enrollment)
df$month = as.factor(ifelse(df$month == 1, "Jan", ifelse(df$month == 2, "Feb", ifelse(df$month == 3, "Mar", ifelse(df$month == 4, "Apr", ifelse(df$month == 5, "May", ifelse(df$month == 6, "Jun", ifelse(df$month == 7, "Jul", ifelse(df$month == 8, "Aug", ifelse(df$month == 9, "Sep", ifelse(df$month == 10, "Oct", ifelse(df$month == 11, "Nov", "Dec"))))))))))))
```

## I use one hot encoding to create dummy variable for the categorical valiables and finally turn is_online into a factor.
```{r dummies}
dmy = dummyVars(" ~ .", data = df)
df = data.frame(predict(dmy, newdata = df))
```

## Summary of Data after Preparation
```{r split_data, echo = FALSE}
#your individual ids; 
#make sure to change the studentid in the SQL-query
testing = df[is.na(df[,"is_online"]),"receipt_id"] 
#provide summary, without the unseen data
df_pred = df[is.na(df[,"is_online"]),]
df = df[!is.na(df[,"is_online"]),]
df$is_online = as.factor(ifelse(df$is_online == 1, "yes", "no"))
# summary(df)
```

# Modelling
# Splitting the dataset for training and validation.
```{r split_train}
set.seed(7313)
smp_siz = round(0.7*nrow(df))
train_row = sample(seq_len(nrow(df)),size = smp_siz)
train = df[train_row,]
test = df[-train_row,]
```

#Model Evaluation: For modeling, we tested: logistic regression, ridge & lasso logistic regression, stepwise logistic regression, decision trees, random forest, and boosted trees (XGBoost). I selected AUC and accuracy as my primary performance metrics to evaluate the model performance since I wanted to obtain a model with an overall good performance that reduces both FNs and FPs. Based on these 2 metrics, the simple logistic regression with the 26 predictors was among the best performing models (Accuracy: 92.07%; AUC: 0.97). This compares against the baseline case of just predicting is_online by picking the mode 'no', which would have 92.66% of being correct. So, the logistic regressions performs only marginally better. In addition, I used the same dataset to train and tune a random forest model and achieved a slightly better performance (+0.28% in accuracy). In this business case, I want to predict whether an transaction is online, which is why the interpretability of the model is slightly more important to weight than performance. Therefore, taking into account the performance-interpretability-tradeoff and Occam’s Razor principle, I opted for the simple logistic regression model. Lastly, the training and test performance differed only marginally, pointing to a good bias-variance-tradeoff.

# Training a logistic regression
```{r logistic_regression, message = FALSE, warning = FALSE}
train.control = trainControl(method = "repeatedcv", number = 10, repeats = 10, verboseIter = F, classProbs = T, summaryFunction = prSummary)

set.seed(7313)
logreg = train(is_online ~ . - receipt_id, data = train, method = "glm", metric = "Recall", trControl = train.control)

# Training performance
# logreg
# summary(logreg)
# confusionMatrix(logreg)
```

# Making predictions on the test datasets
```{r predictions, message = FALSE, warning = FALSE}
test.predictions = data.frame(pred = predict(logreg, newdata = test), obs = test$is_online)

prob.predictions = predict(logreg, newdata = test, type = "prob")

test.predictions = bind_cols(test.predictions, prob.predictions)
test.predictions = test.predictions %>% mutate(obs = as.factor(obs)) %>% mutate(obs = factor(obs, levels = c("no", "yes")))

confusionMatrix(test.predictions$pred, test.predictions$obs)
twoClassSummary(test.predictions, lev = c("no", "yes"))
prSummary(test.predictions, lev = c("no", "yes"))
```

```{r, include = FALSE}
pred = data.frame(receipt_id = testing, pred = predict(logreg, newdata = df_pred))
pred$pred = as.numeric(pred$pred)
pred$pred = pred$pred - 1
n = 42205
write.csv(pred, paste(n,".csv",""))
```

**Submit the knitted pdf and the csv on Canvas**
